---
title: "Invertebrates BC - Model comparison for acc rain"
author: "Dr. Maja Ilić"
date: "3/24/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Info  

This script fits linear models of the invertebrates resilience measure with the following random effect structures (as discussed with PC, 23.03.2020):   

1. (1|day) + (1|Farm) + (acc_rain|Region)    
2. (1|day) + (acc_rain|Region)      
3. (1|Farm) + (acc_rain|Region)  
4. (1|day) + (1|Region/Farm)   
5. (1|day)  
6. (1|Region/Farm)  
7. (1|day) + (1|Region/Farm) + (acc_rain|Region)   
8. (1|Region/Farm) + (acc_rain|Region)   
9. (1|day) + (1|Region/Farm) + (acc_rain - 1|Region)   
10. (1|Region/Farm) + (acc_rain - 1|Region)  

Fixed effect structure in these "full models" is given as: **day * acc_rain**        

The response is the invertebrate Bray-Curtis Dissimilarity Index, calculated by comparing the invertebrate community in each plot with the average invertebrate community of the control plots within the same site and time step (reference community). BC-Index of 0 indicates that the two compared communities are equal and share all species, while BC-Index of 1 indicates that the two compared communities do not share any species.      

**NOTE:** First 6 random effect structures were suggested by PC, while the last 4 were added by MI (out of curiousity).  

**NOTE 2:** Both day and acc_rain were rescale to be able to fit the models.  

## Set working directory  

```{r}
# Maja's desktop PC

setwd("C:/Users/3054311/Documents/My Documents/Grassland project/05_Resilience/Model comparison")
```

## Packages  

```{r warning = FALSE, message = FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
library(car)
library(performance)
library(insight)
library(see)
library(effects)
library(afex)
library(optimx)
library(nloptr)
library(dfoptim)
library(pander)
library(fmsb)
library(RColorBrewer)
library(scales)
```

## Load invertebrates BC and rainfall data  

```{r}
load(file = "C:/Users/3054311/Documents/My Documents/Grassland project/05_Resilience/Data/Resilience_Inverts_with_Rainfall.R")
names(resilience_inverts)[names(resilience_inverts) == "rain_out"] <- "acc_rain"
```

## Rescaling day and acc_rain  

```{r}
resilience_inverts_sc <- resilience_inverts
resilience_inverts_sc$day <- scale(resilience_inverts_sc$day)
resilience_inverts_sc$acc_rain <- scale(resilience_inverts_sc$acc_rain)
```

# Model 1: (1|day) + (1|Farm) + (acc_rain|Region)  

```{r}
rand1 <- lmer(inverts_BC_all ~ day * acc_rain + (1|day) + (1|Farm) + (acc_rain|Region),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand1)

theta <- getME(rand1,"theta")
diag.element <- getME(rand1,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand1)
check_convergence(rand1)
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand1)
```

## SD and correlation of the random effects:  

```{r}
summary(rand1)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand1)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand1), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand1)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand1)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out1 <- data.frame(resilience_inverts, "Fitted" = fitted(rand1))  

ggplot(out1, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out1, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 2: (1|day) + (acc_rain|Region)  

```{r}
rand2 <- lmer(inverts_BC_all ~ day * acc_rain + (1|day) + (acc_rain|Region),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand2)

theta <- getME(rand2,"theta")
diag.element <- getME(rand2,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand2)
check_convergence(rand2)
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand2)
```

## SD and correlation of the random effects:  

```{r}
summary(rand2)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand2)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand2), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand2)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand2)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out2 <- data.frame(resilience_inverts, "Fitted" = fitted(rand2))  

ggplot(out2, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out2, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 3: (1|Farm) + (acc_rain|Region)  

```{r}
rand3 <- lmer(inverts_BC_all ~ day * acc_rain + (1|Farm) + (acc_rain|Region),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand3)

theta <- getME(rand3,"theta")
diag.element <- getME(rand3,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand3)
check_convergence(rand3)
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand3)
```

## SD and correlation of the random effects:  

```{r}
summary(rand3)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand3)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand3), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand3)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand3)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out3 <- data.frame(resilience_inverts, "Fitted" = fitted(rand3))  

ggplot(out3, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out3, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 4: (1|day) + (1|Region/Farm)  

```{r}
rand4 <- lmer(inverts_BC_all ~ day * acc_rain + (1|day) + (1|Region/Farm),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand4)

theta <- getME(rand4,"theta")
diag.element <- getME(rand4,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand4)
check_convergence(rand4)
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand4)
```

## SD and correlation of the random effects:  

```{r}
summary(rand4)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand4)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand4), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand4)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand4)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out4 <- data.frame(resilience_inverts, "Fitted" = fitted(rand4))  

ggplot(out4, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out4, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 5: (1|day)  

```{r}
rand5 <- lmer(inverts_BC_all ~ day * acc_rain + (1|day),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand5)

theta <- getME(rand5,"theta")
diag.element <- getME(rand5,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand5)
check_convergence(rand5)
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand5)
```

## SD and correlation of the random effects:  

```{r}
summary(rand5)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand5)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand5), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand5)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand5)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out5 <- data.frame(resilience_inverts, "Fitted" = fitted(rand5))  

ggplot(out5, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out5, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 6: (1|Region/Farm)  

```{r}
rand6 <- lmer(inverts_BC_all ~ day * acc_rain + (1|Region/Farm),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand6)

theta <- getME(rand6,"theta")
diag.element <- getME(rand6,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand6)
check_convergence(rand6)
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand6)
```

## SD and correlation of the random effects:  

```{r}
summary(rand6)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand6)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand6), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand6)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand6)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out6 <- data.frame(resilience_inverts, "Fitted" = fitted(rand6))  

ggplot(out6, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out6, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 7: (1|day) + (1|Region/Farm) + (acc_rain|Region)  

```{r}
rand7 <- lmer(inverts_BC_all ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain|Region),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand7)

theta <- getME(rand7,"theta")
diag.element <- getME(rand7,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand7)
check_convergence(rand7)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand7)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand7_bobyqa <- lmer(inverts_BC_all ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain|Region),
                     data = resilience_inverts_sc,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand7_nlminb <- lmer(inverts_BC_all ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain|Region),
                     data = resilience_inverts_sc,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand7_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand7_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand7_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand7_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand7_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand7_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out7 <- data.frame(resilience_inverts, "Fitted" = fitted(rand7_nlminb))  

ggplot(out7, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out7, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 8: (1|Region/Farm) + (acc_rain|Region)  

```{r}
rand8 <- lmer(inverts_BC_all ~ day * acc_rain + (1|Region/Farm) + (acc_rain|Region),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand8)

theta <- getME(rand8,"theta")
diag.element <- getME(rand8,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand8)
check_convergence(rand8)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand8)
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand8_nlminb <- lmer(inverts_BC_all ~ day * acc_rain + (1|Region/Farm) + (acc_rain|Region),
                     data = resilience_inverts_sc,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand8_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand8_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand8_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand8_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand8_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand8_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out8 <- data.frame(resilience_inverts, "Fitted" = fitted(rand8_nlminb))  

ggplot(out8, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|Region/Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out8, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|Region/Farm) + (acc_rain|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 9: (1|day) + (1|Region/Farm) + (acc_rain - 1|Region)  

```{r}
rand9 <- lmer(inverts_BC_all ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain - 1|Region),
              data = resilience_inverts_sc,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand9)

theta <- getME(rand9,"theta")
diag.element <- getME(rand9,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand9)
check_convergence(rand9)
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand9)
```

## SD and correlation of the random effects:  

```{r}
summary(rand9)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand9)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand9), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand9)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand9)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out9 <- data.frame(resilience_inverts, "Fitted" = fitted(rand9))  

ggplot(out9, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain - 1|Region)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out9, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain - 1|Region)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 10: (1|Region/Farm) + (acc_rain - 1|Region)  

```{r}
rand10 <- lmer(inverts_BC_all ~ day * acc_rain + (1|Region/Farm) + (acc_rain - 1|Region),
               data = resilience_inverts_sc,
               REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand10)

theta <- getME(rand10,"theta")
diag.element <- getME(rand10,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand10)
check_convergence(rand10)
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand10)
```

## SD and correlation of the random effects:  

```{r}
summary(rand10)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand10)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand10), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand10)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand10)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out10 <- data.frame(resilience_inverts, "Fitted" = fitted(rand10))  

ggplot(out10, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|Region/Farm) + (acc_rain - 1|Region)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out10, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.2,0.6)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "invertebrates_BC ~ day * acc_rain + (1|day) + (1|Region/Farm) + (acc_rain - 1|Region)",
       subtitle = "lm(fitted data ~ day)")
```

# Model comparison  

```{r}
result <- compare_performance(rand1,rand2,rand3,
                              rand4,rand5,rand6,
                              rand7_nlminb,rand8_nlminb,
                              rand9,rand10)
```

## Plot results using package fmsb  

```{r warning = FALSE, message = FALSE, fig.width = 9, fig.height = 5}
mod_comp <- as.data.frame(result[,c(3,4,6,7,9)])
rownames(mod_comp) <- result$Model
colnames(mod_comp)[which(colnames(mod_comp) == "R2_conditional")] <- "R2_cond"
colnames(mod_comp)[which(colnames(mod_comp) == "R2_marginal")] <- "R2_marg"

# Print results

pander(mod_comp, style = 'rmarkdown')

# Change NAs to zero, and multiply AIC and BIC by -1 (for visualization purposes)

mod_comp[is.na(mod_comp)] <- 0 
mod_comp$AIC <- mod_comp$AIC * -1
mod_comp$BIC <- mod_comp$BIC * -1

# Add max and min

mod_comp <- rbind(c(max(mod_comp$AIC),max(mod_comp$BIC),max(mod_comp$R2_cond),
                    max(mod_comp$R2_marg),max(mod_comp$BF)),
                  c(min(mod_comp$AIC),min(mod_comp$BIC),min(mod_comp$R2_cond),
                    min(mod_comp$R2_marg),min(mod_comp$BF)),
                  mod_comp)

# Set graphic colors 

colors_border <- colorRampPalette(brewer.pal(8, "Spectral"))(nrow(mod_comp)-2)
colors_in <- alpha(colors_border,0.2)

# Plot results 

par(mfrow = c(1,2), mar = c(1,0,2,0))

radarchart(mod_comp, axistype = 1,
           pcol = colors_border , pfcol = colors_in , plwd = 2 , plty = 1,
           cglcol = "grey", cglty = 2, axislabcol = "grey", 
           vlcex = 1.4, calcex = 1.2
)

# Add a legend

plot(NULL,xlim = c(0,1), ylim = c(0,1), axes = F)

legend(x = 0.1, y = 0.5, legend = rownames(mod_comp)[-c(1,2)], bty = "n", pch = 0, 
       col = colors_border, text.col = "black", cex = 1.4, pt.cex = 3, pt.lwd = 2, yjust = 0.5)
```

## Plot results for each model separately  

```{r fig.width = 9, fig.height = 9}
par(mfrow = c(4,3), mar = c(1,2,2,1))

for (i in 1:(nrow(mod_comp)-2)){
  radarchart(mod_comp[c(1,2,(i+2)),], axistype = 1,
             pcol = colors_border[i] , pfcol = colors_in[i] , plwd = 2 , plty = 1,
             cglcol = "grey90", cglty = 2, axislabcol = "grey60", calcex = 1,
             vlcex = 1, 
             title = rownames(mod_comp)[i+2], cex.main = 1.2
  )
}
```
