---
title: "Plant biomass ratio - Model comparison for treatment"
author: "Dr. Maja Ilić"
date: "3/23/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Info  

This script fits linear models of the plant biomass resilience measure with the following random effect structures (as discussed with PC, 23.03.2020):    

1. (1|day) + (1|Farm) + (treatment|Region)    
2. (1|day) + (treatment|Region)      
3. (1|Farm) + (treatment|Region)  
4. (1|day) + (1|Region/Farm)   
5. (1|day)  
6. (1|Region/Farm)  
7. (1|day) + (1|Region/Farm) + (treatment|Region)   
8. (1|Region/Farm) + (treatment|Region)   
9. (1|day) + (1|Region/Farm) + (treatment - 1|Region)   
10. (1|Region/Farm) + (treatment - 1|Region)  

Fixed effect structure in these "full models" is given as: **day * treatment**        

The response is the ratio of dry weight from non repeated measures in each plot to the average of control plots at the same site (farm) and time step.    

**NOTE:** First 6 random effect structures were suggested by PC, while the last 4 were added by MI (out of curiousity).     

## Set working directory  

```{r}
# Maja's desktop PC

setwd("C:/Users/3054311/Documents/My Documents/Grassland project/05_Resilience/Model comparison")
```

## Packages  

```{r warning = FALSE, message = FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
library(car)
library(performance)
library(insight)
library(see)
library(effects)
library(afex)
library(optimx)
library(nloptr)
library(dfoptim)
library(pander)
library(fmsb)
library(RColorBrewer)
library(scales)
```

## Load plant biomass ratios and rainfall data  

```{r}
load(file = "C:/Users/3054311/Documents/My Documents/Grassland project/05_Resilience/Data/Resilience_Plant_with_Rainfall_non_rm.R")
```

# Model 1: (1|day) + (1|Farm) + (treatment|Region)  

```{r}
rand1 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand1)

theta <- getME(rand1,"theta")
diag.element <- getME(rand1,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand1)
check_convergence(rand1)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand1)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand1_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand1_bobyqa)
```

## SD and correlation of the random effects:  

```{r}
summary(rand1_bobyqa)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand1_bobyqa)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand1_bobyqa), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand1_bobyqa)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand1_bobyqa)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out1 <- data.frame(resilience_plant, "Fitted" = fitted(rand1_bobyqa))  

ggplot(out1, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: bobyqa")

ggplot(out1, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: bobyqa")
```

# Model 2: (1|day) + (treatment|Region)  

```{r}
rand2 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (treatment|Region),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand2)

theta <- getME(rand2,"theta")
diag.element <- getME(rand2,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand2)
check_convergence(rand2)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand2)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand2_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (treatment|Region),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nmkbw:  

```{r}
rand2_nmkb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (treatment|Region),
                   data = resilience_plant,
                   REML = T,
                   control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                         optCtrl = list(method = "nmkb", 
                                                        starttests = FALSE, 
                                                        kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand2_nmkb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand2_nmkb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand2_nmkb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand2_nmkb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand2_nmkb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand2_nmkb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out2 <- data.frame(resilience_plant, "Fitted" = fitted(rand2_nmkb))  

ggplot(out2, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")

ggplot(out2, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")
```

# Model 3: (1|Farm) + (treatment|Region)  

```{r}
rand3 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Farm) + (treatment|Region),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand3)

theta <- getME(rand3,"theta")
diag.element <- getME(rand3,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand3)
check_convergence(rand3)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand3)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand3_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|Farm) + (treatment|Region),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nmkb:  

```{r}
rand3_nmkb <- lmer(plot_meanC_ratio ~ day * treatment + (1|Farm) + (treatment|Region),
                   data = resilience_plant,
                   REML = T,
                   control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                         optCtrl = list(method = "nmkb", 
                                                        starttests = FALSE, 
                                                        kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand3_nmkb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand3_nmkb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand3_nmkb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand3_nmkb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand3_nmkb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand3_nmkb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out3 <- data.frame(resilience_plant, "Fitted" = fitted(rand3_nmkb))  

ggplot(out3, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")

ggplot(out3, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")
```

# Model 4: (1|day) + (1|Region/Farm)  

```{r}
rand4 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand4)

theta <- getME(rand4,"theta")
diag.element <- getME(rand4,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand4)
check_convergence(rand4)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand4)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand4_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa",
                                           optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nmkb:  

```{r}
rand4_nmkb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm),
                   data = resilience_plant,
                   REML = T,
                   control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                         optCtrl = list(method = "nmkb", 
                                                        starttests = FALSE, 
                                                        kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand4_nmkb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand4_nmkb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand4_nmkb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand4_nmkb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand4_nmkb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand4_nmkb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out4 <- data.frame(resilience_plant, "Fitted" = fitted(rand4_nmkb))  

ggplot(out4, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")

ggplot(out4, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")
```

# Model 5: (1|day)  

```{r}
rand5 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand5)

theta <- getME(rand5,"theta")
diag.element <- getME(rand5,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand5)
check_convergence(rand5)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand5)
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand5_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand5_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand5_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand5_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand5_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand5_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand5_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out5 <- data.frame(resilience_plant, "Fitted" = fitted(rand5_nlminb))  

ggplot(out5, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out5, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 6: (1|Region/Farm)  

```{r}
rand6 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand6)

theta <- getME(rand6,"theta")
diag.element <- getME(rand6,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand6)
check_convergence(rand6)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand6)
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand6_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand6_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand6_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand6_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand6_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand6_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand6_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out6 <- data.frame(resilience_plant, "Fitted" = fitted(rand6_nlminb))  

ggplot(out6, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out6, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 7: (1|day) + (1|Region/Farm) + (treatment|Region)  

```{r}
rand7 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand7)

theta <- getME(rand7,"theta")
diag.element <- getME(rand7,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand7)
check_convergence(rand7)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand7)
```

## Refitting the model with the optimizer Nelder_Mead:  

```{r}
rand7_NM <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region),
                 data = resilience_plant,
                 REML = T,
                 control = lmerControl(optimizer = "Nelder_Mead", 
                                       optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand7_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand7_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand7_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand7_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand7_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand7_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand7_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out7 <- data.frame(resilience_plant, "Fitted" = fitted(rand7_nlminb))  

ggplot(out7, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out7, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 8: (1|Region/Farm) + (treatment|Region)  

```{r}
rand8 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand8)

theta <- getME(rand8,"theta")
diag.element <- getME(rand8,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand8)
check_convergence(rand8)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand8)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand8_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand8_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand8_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand8_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand8_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand8_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand8_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand8_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out8 <- data.frame(resilience_plant, "Fitted" = fitted(rand8_nlminb))  

ggplot(out8, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out8, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 9: (1|day) + (1|Region/Farm) + (treatment - 1|Region)  

```{r}
rand9 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region),
              data = resilience_plant,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand9)

theta <- getME(rand9,"theta")
diag.element <- getME(rand9,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand9)
check_convergence(rand9)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand9)
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand9_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region),
                     data = resilience_plant,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand9_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand9_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand9_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand9_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand9_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand9_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out9 <- data.frame(resilience_plant, "Fitted" = fitted(rand9_nlminb))  

ggplot(out9, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out9, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 10: (1|Region/Farm) + (treatment - 1|Region)  

```{r}
rand10 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment - 1|Region),
               data = resilience_plant,
               REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand10)

theta <- getME(rand10,"theta")
diag.element <- getME(rand10,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand10)
check_convergence(rand10)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand10)
```

## Refitting the model with the optimizer nmkb:  

```{r}
rand10_nmkb <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment - 1|Region),
                    data = resilience_plant,
                    REML = T,
                    control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                          optCtrl = list(method = "nmkb", 
                                                         starttests = FALSE, 
                                                         kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand10_nmkb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand10_nmkb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand10_nmkb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand10_nmkb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand10_nmkb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand10_nmkb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out10 <- data.frame(resilience_plant, "Fitted" = fitted(rand10_nmkb))  

ggplot(out10, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")

ggplot(out10, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.5,1.1)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "plant_biomass_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nmkb")
```

# Model comparison  

```{r}
result <- compare_performance(rand1_bobyqa,rand2_nmkb,rand3_nmkb,
                              rand4_nmkb,rand5_nlminb,rand6_nlminb,
                              rand7_nlminb,rand8_nlminb,
                              rand9_nlminb,rand10_nmkb)
```

## Plot results using package fmsb  

```{r warning = FALSE, message = FALSE, fig.width = 9, fig.height = 5}
mod_comp <- as.data.frame(result[,c(3,4,6,9)])
rownames(mod_comp) <- result$Model
colnames(mod_comp)[which(colnames(mod_comp) == "R2_marginal")] <- "R2_marg"

# Print results

pander(mod_comp, style = 'rmarkdown')

# Change NAs to zero, and multiply AIC and BIC by -1 (for visualization purposes)

mod_comp[is.na(mod_comp)] <- 0 
mod_comp$AIC <- mod_comp$AIC * -1
mod_comp$BIC <- mod_comp$BIC * -1

# Add max and min

mod_comp <- rbind(c(max(mod_comp$AIC),max(mod_comp$BIC),
                    max(mod_comp$R2_marg),max(mod_comp$BF)),
                  c(min(mod_comp$AIC),min(mod_comp$BIC),
                    min(mod_comp$R2_marg),min(mod_comp$BF)),
                  mod_comp)

# Set graphic colors 

colors_border <- colorRampPalette(brewer.pal(8, "Spectral"))(nrow(mod_comp)-2)
colors_in <- alpha(colors_border,0.2)

# Plot results 

par(mfrow = c(1,2), mar = c(1,0,2,0))

radarchart(mod_comp, axistype = 1,
           pcol = colors_border , pfcol = colors_in , plwd = 2 , plty = 1,
           cglcol = "grey", cglty = 2, axislabcol = "grey", 
           vlcex = 1.4, calcex = 1.2
)

# Add a legend

plot(NULL,xlim = c(0,1), ylim = c(0,1), axes = F)

legend(x = 0.1, y = 0.5, legend = rownames(mod_comp)[-c(1,2)], bty = "n", pch = 0, 
       col = colors_border, text.col = "black", cex = 1.4, pt.cex = 3, pt.lwd = 2, yjust = 0.5)
```

## Plot results for each model separately  

```{r fig.width = 9, fig.height = 9}
par(mfrow = c(4,3), mar = c(1,2,2,1))

for (i in 1:(nrow(mod_comp)-2)){
  radarchart(mod_comp[c(1,2,(i+2)),], axistype = 1,
             pcol = colors_border[i] , pfcol = colors_in[i] , plwd = 2 , plty = 1,
             cglcol = "grey90", cglty = 2, axislabcol = "grey60", calcex = 1,
             vlcex = 1, 
             title = rownames(mod_comp)[i+2], cex.main = 1.2
  )
}
```
