---
title: "Soil moisture ratio rm - Model comparison for treatment"
author: "Dr. Maja Ilić"
date: "4/22/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Info  

This script fits linear models of the soil moisture resilience measure (rm) with the following random effect structures (as discussed with PC, 23.03.2020):    

1. (1|day) + (1|Farm) + (treatment|Region)    
2. (1|day) + (treatment|Region)      
3. (1|Farm) + (treatment|Region)  
4. (1|day) + (1|Region/Farm)   
5. (1|day)  
6. (1|Region/Farm)  
7. (1|day) + (1|Region/Farm) + (treatment|Region)   
8. (1|Region/Farm) + (treatment|Region)   
9. (1|day) + (1|Region/Farm) + (treatment - 1|Region)   
10. (1|Region/Farm) + (treatment - 1|Region)  

Fixed effect structure in these "full models" is given as: **day * treatment**        

The response is the ratio of soil moisture in each plot to the average soil moisture of control plots at the same site (farm) and time step. The calculated ratio corresponds to the repeated measure of plant biomass (starting at middrought).    

**NOTE:** First 6 random effect structures were suggested by PC, while the last 4 were added by MI (out of curiousity).     

## Set working directory  

```{r}
# Maja's desktop PC

setwd("C:/Users/3054311/Documents/My Documents/Grassland project/05_Resilience/Model comparison")
```

## Packages  

```{r warning = FALSE, message = FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
library(car)
library(performance)
library(insight)
library(see)
library(effects)
library(afex)
library(optimx)
library(nloptr)
library(dfoptim)
library(pander)
library(fmsb)
library(RColorBrewer)
library(scales)
```

## Load soil moisture ratios and rainfall data  

```{r}
load(file = "C:/Users/3054311/Documents/My Documents/Grassland project/05_Resilience/Data/Resilience_Soil_with_Rainfall_rm.RData")

resilience_soil_rm <- resilience_soil_rm[!is.na(resilience_soil_rm$plot_meanC_ratio),]
```

# Model 1: (1|day) + (1|Farm) + (treatment|Region)  

```{r}
rand1 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand1)

theta <- getME(rand1,"theta")
diag.element <- getME(rand1,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand1)
check_convergence(rand1)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand1)
```

## Refitting the model with the optimizer nlminb  

```{r}
rand1_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand1_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand1_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand1_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand1_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand1_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand1_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out1 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand1_nlminb))  

ggplot(out1, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out1, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 2: (1|day) + (treatment|Region)  

```{r}
rand2 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (treatment|Region),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand2)

theta <- getME(rand2,"theta")
diag.element <- getME(rand2,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand2)
check_convergence(rand2)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand2)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand2_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (treatment|Region),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand2_bobyqa)
```

## SD and correlation of the random effects:  

```{r}
summary(rand2_bobyqa)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand2_bobyqa)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand2_bobyqa), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand2_bobyqa)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand2_bobyqa)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out2 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand2_bobyqa))  

ggplot(out2, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: bobyqa")

ggplot(out2, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: bobyqa")
```

# Model 3: (1|Farm) + (treatment|Region)  

```{r}
rand3 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Farm) + (treatment|Region),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand3)

theta <- getME(rand3,"theta")
diag.element <- getME(rand3,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand3)
check_convergence(rand3)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand3)
```

## Refitting the model with the optimizer Nelder_Mead:  

```{r}
rand3_NM <- lmer(plot_meanC_ratio ~ day * treatment + (1|Farm) + (treatment|Region),
                 data = resilience_soil_rm,
                 REML = T,
                 control = lmerControl(optimizer = "Nelder_Mead", 
                                       optCtrl = list(maxfun = 2e5)))
```

## Model validation:  

```{r fig.width = 12, fig.height = 9, warning = F, message = F}
check_model(rand3_NM)
```

## SD and correlation of the random effects:  

```{r}
summary(rand3_NM)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand3_NM)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand3_NM), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand3_NM)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand3_NM)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out3 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand3_NM))  

ggplot(out3, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: Nelder_Mead")

ggplot(out3, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: Nelder_Mead")
```

# Model 4: (1|day) + (1|Region/Farm)  

```{r}
rand4 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand4)

theta <- getME(rand4,"theta")
diag.element <- getME(rand4,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand4)
check_convergence(rand4)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand4)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand4_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (Region/Farm),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa",
                                           optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand4_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (Region/Farm),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

Although the model was refitted with all available optimizers (not included in this report), I could not get rid of the warnings. Therefore, I will keep the inital full model.

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand4)
```

## SD and correlation of the random effects:  

```{r}
summary(rand4)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand4)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand4), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand4)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand4)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out4 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand4))  

ggplot(out4, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out4, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 5: (1|day)  

```{r}
rand5 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand5)

theta <- getME(rand5,"theta")
diag.element <- getME(rand5,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand5)
check_convergence(rand5)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand5)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand5_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|day),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa",
                                           optCtrl = list(maxfun = 2e5)))
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand5_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand5_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand5_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand5_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand5_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand5_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand5_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out5 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand5_nlminb))  

ggplot(out5, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out5, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 6: (1|Region/Farm)  

```{r}
rand6 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand6)

theta <- getME(rand6,"theta")
diag.element <- getME(rand6,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand6)
check_convergence(rand6)
```

## Model validation:  

```{r fig.width = 9, fig.height = 9, warning = F, message = F}
check_model(rand6)
```

## SD and correlation of the random effects:  

```{r}
summary(rand6)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand6)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand6), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand6)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand6)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out6 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand6))  

ggplot(out6, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")

ggplot(out6, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|Region/Farm)",
       subtitle = "lm(fitted data ~ day)")
```

# Model 7: (1|day) + (1|Region/Farm) + (treatment|Region)  

```{r}
rand7 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand7)

theta <- getME(rand7,"theta")
diag.element <- getME(rand7,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand7)
check_convergence(rand7)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand7)
```

## Refitting the model with the optimizer bobyqa:  

```{r}
rand7_bobyqa <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
```

Refitting the model with the optimizer bobyqa solved the convergence issue, but the model is still singular.  

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand7_bobyqa)
```

## SD and correlation of the random effects:  

```{r}
summary(rand7_bobyqa)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand7_bobyqa)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand7_bobyqa), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand7_bobyqa)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand7_bobyqa)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out7 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand7_bobyqa))  

ggplot(out7, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: bobyqa")

ggplot(out7, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: bobyqa")
```

# Model 8: (1|Region/Farm) + (treatment|Region)  

```{r}
rand8 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand8)

theta <- getME(rand8,"theta")
diag.element <- getME(rand8,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand8)
check_convergence(rand8)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand8)
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand8_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand8_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand8_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand8_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand8_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand8_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand8_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out8 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand8_nlminb))  

ggplot(out8, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out8, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|Region/Farm) + (treatment|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 9: (1|day) + (1|Region/Farm) + (treatment - 1|Region)  

```{r}
rand9 <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region),
              data = resilience_soil_rm,
              REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand9)

theta <- getME(rand9,"theta")
diag.element <- getME(rand9,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand9)
check_convergence(rand9)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand9)
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand9_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region),
                     data = resilience_soil_rm,
                     REML = T,
                     control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", 
                                                          starttests = FALSE, 
                                                          kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand9_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand9_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand9_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand9_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand9_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand9_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out9 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand9_nlminb))  

ggplot(out9, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out9, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model 10: (1|Region/Farm) + (treatment - 1|Region)  

```{r}
rand10 <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment - 1|Region),
               data = resilience_soil_rm,
               REML = T) 
```

## Check the model for singularity and convergence:  

```{r}
isSingular(rand10)

theta <- getME(rand10,"theta")
diag.element <- getME(rand10,"lower") == 0
any(theta[diag.element] < 1e-5)

check_singularity(rand10)
check_convergence(rand10)
```

## Refitting the model with different optimizers:  

```{r}
allFit(rand10)
```

## Refitting the model with the optimizer nlminb:  

```{r}
rand10_nlminb <- lmer(plot_meanC_ratio ~ day * treatment + (1|Region/Farm) + (treatment - 1|Region),
                      data = resilience_soil_rm,
                      REML = T,
                      control = lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                            optCtrl = list(method = "nlminb", 
                                                           starttests = FALSE, 
                                                           kkt = FALSE)))
```

## Model validation:  

```{r fig.width = 7, fig.height = 8, warning = F, message = F}
check_model(rand10_nlminb)
```

## SD and correlation of the random effects:  

```{r}
summary(rand10_nlminb)$varcor
```

## Variance of the random effects:  

```{r warning = F}
pander(summary(rand10_nlminb)$varcor, style = 'rmarkdown')
```

## Estimates for the random effects:  

```{r warning = F}
pander(ranef(rand10_nlminb), style = 'rmarkdown')
```

## Significance of the random effects:  

```{r message = F}
pander(rand(rand10_nlminb)[c(2,3,5,6)], style = 'rmarkdown')
```

## Significance of the fixed effects:  

```{r}
pander(anova(rand10_nlminb)[-3], style = 'rmarkdown')
```

## Visualize the model output using fitted values:  

```{r fig.width = 8, fig.height = 7}
out10 <- data.frame(resilience_soil_rm, "Fitted" = fitted(rand10_nlminb))  

ggplot(out10, 
       aes(x = day, y = Fitted)) +
  facet_wrap(~ Region) +
  geom_smooth(aes(color = treatment, linetype = Farm),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_fill_manual(values = c("blue","red")) +
  scale_color_manual(values = c("blue","red")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")

ggplot(out10, 
       aes(x = day, y = Fitted)) +
  geom_smooth(aes(color = Region, linetype = treatment),
              method = "lm", se = F, size = 0.8) +
  theme_minimal() +
  theme(strip.background = element_rect(),
        panel.spacing = unit(1, "lines"),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(size = 11, face = "bold", margin = margin(15,0,0,0)),
        axis.title.y = element_text(size = 11, face = "bold", margin = margin(0,15,0,0)),
        plot.title = element_text(size = 12, face = "bold")) +
  scale_y_continuous(limits = c(0.4,1.3)) +
  scale_x_continuous(limits = c(0,64), breaks = c(0,8,16,32,64)) +
  labs(x = "Days since the end of the simulated drought",
       title = "soil_moisture_ratio ~ day * treatment + (1|day) + (1|Region/Farm) + (treatment - 1|Region)",
       subtitle = "lm(fitted data ~ day), optimizer: nlminb")
```

# Model comparison  

```{r}
result <- compare_performance(rand1_nlminb, rand2_bobyqa, rand3_NM,
                              rand4, rand5_nlminb, rand6,
                              rand7_bobyqa, rand8_nlminb,
                              rand9_nlminb, rand10_nlminb)
```

## Plot results using package fmsb  

```{r warning = FALSE, message = FALSE, fig.width = 9, fig.height = 5}
mod_comp <- as.data.frame(result[,c(3,4,5,6,9)])
rownames(mod_comp) <- result$Model
colnames(mod_comp)[which(colnames(mod_comp) == "R2_conditional")] <- "R2_cond"
colnames(mod_comp)[which(colnames(mod_comp) == "R2_marginal")] <- "R2_marg"

# Print results

pander(mod_comp, style = 'rmarkdown')

# Change NAs to zero, and multiply AIC and BIC by -1 (for visualization purposes)

mod_comp[is.na(mod_comp)] <- 0 
mod_comp$AIC <- mod_comp$AIC * -1
mod_comp$BIC <- mod_comp$BIC * -1

# Add max and min

mod_comp <- rbind(c(max(mod_comp$AIC),max(mod_comp$BIC),max(mod_comp$R2_cond),
                    max(mod_comp$R2_marg),max(mod_comp$BF)),
                  c(min(mod_comp$AIC),min(mod_comp$BIC),min(mod_comp$R2_cond),
                    min(mod_comp$R2_marg),min(mod_comp$BF)),
                  mod_comp)

# Set graphic colors 

colors_border <- colorRampPalette(brewer.pal(8, "Spectral"))(nrow(mod_comp)-2)
colors_in <- alpha(colors_border,0.2)

# Plot results 

par(mfrow = c(1,2), mar = c(1,0,2,0))

radarchart(mod_comp, axistype = 1,
           pcol = colors_border , pfcol = colors_in , plwd = 2 , plty = 1,
           cglcol = "grey", cglty = 2, axislabcol = "grey", 
           vlcex = 1.4, calcex = 1.2
)

# Add a legend

plot(NULL,xlim = c(0,1), ylim = c(0,1), axes = F)

legend(x = 0.1, y = 0.5, legend = rownames(mod_comp)[-c(1,2)], bty = "n", pch = 0, 
       col = colors_border, text.col = "black", cex = 1.4, pt.cex = 3, pt.lwd = 2, yjust = 0.5)
```

## Plot results for each model separately  

```{r fig.width = 9, fig.height = 9}
par(mfrow = c(4,3), mar = c(1,2,2,1))

for (i in 1:(nrow(mod_comp)-2)){
  radarchart(mod_comp[c(1,2,(i+2)),], axistype = 1,
             pcol = colors_border[i] , pfcol = colors_in[i] , plwd = 2 , plty = 1,
             cglcol = "grey90", cglty = 2, axislabcol = "grey60", calcex = 1,
             vlcex = 1, 
             title = rownames(mod_comp)[i+2], cex.main = 1.2
  )
}
```
